### 2016-09-21 paper.tex

\documentclass[aps,rmp,twocolumn,amsmath,amssymb,nofootinbib,superscriptaddress]{revtex4}

\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\op}[2]{\hat{\textbf{#1}}_{#2}}
\newcommand{\dagop}[2]{\hat{\textbf{#1}}_{#2}^\dag}
\usepackage[pdftex]{graphicx}
\usepackage{mathrsfs}
\usepackage[colorlinks]{hyperref}

\begin{document}

\bibliographystyle{apsrev}

%
% Title
%

\title{Boson-sampling with single photons: A roadmap to post-classical computation}

%
% Authors
%

%\author{Peter P. Rohde}
%\email[]{dr.rohde@gmail.com}
%\homepage{http://www.peterrohde.org}
%\affiliation{Centre for Quantum Computation and Intelligent Systems (QCIS), Faculty of Engineering \& Information Technology, University of Technology Sydney, NSW 2007, Australia}

\date{\today}

\frenchspacing

%
% Abstract
%

\begin{abstract}
\end{abstract}

\maketitle

\tableofcontents

%
% Acknowledgments
%

\begin{acknowledgments}
\end{acknowledgments}

%
% Bibliography
%

\bibliography{bibliography}

\end{document}